---
title: "Predictive Modeling of College Tuition Cost"
author: "by mas16"
date: "25 August 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The College Tuition Crisis

* As of 2019, the collective student debt in the United States surpassed [$1.5 trillion USD](https://www.cnbc.com/2019/05/16/students-families-need-support-systems-to-handle-college-debt-crisis.html)

* Over 2.5 million Americans have student loan debt [exceeding $100,000 USD](https://www.forbes.com/sites/zackfriedman/2019/02/25/student-loan-debt-statistics-2019/#65585571133f).

* [10.5 million students](https://nces.ed.gov/programs/coe/indicator_cha.asp) are projected to enroll in full-time college programs in the United States each year from now until 2028

## Addressing the College Tuition Crisis

* One potential solution to reducing student loan debt is to inform applicants of the high cost of college tuition **before** they enroll

* A web application has been created to predict the cost of college tuition through the year 2035

* The app uses [data](https://nces.ed.gov/fastfacts/display.asp?id=76) from the US Department of Education and linear regression to model the dependence of college tuition cost in USD on year of enrollment

## Predictive Modeling of College Tuition

* The cost of college tuition has [increased linearly]((https://nces.ed.gov/fastfacts/display.asp?id=76)) over the last 20 years

* This data can be fit to a simple linear model to predict the future cost of college tuition

* Example college tuition cost data is shown below for Four Year colleges:

```{r, message = FALSE}
# Read college tuition data
tuition <- read.csv("./tuition.csv", header=TRUE, 
                    colClasses=rep("numeric",8))
# Convert to data frame
tuition <- data.frame(tuition)
# Example Data
head(tuition[,c(1,3)])
```

## Predicting the Cost of College Tuition Using the Web App

The web app can be run by clicking the link below:  
- Shiny App to Precidic College Tuition: [https://mas16.shinyapps.io/Tuition/](https://mas16.shinyapps.io/Tuition/) 

The code for the web app can be accessed in the GitHub repo linked below:  
- GitHub Repo: [https://github.com/mas16/Tuition](https://github.com/mas16/Tuition)

This presentation can be accessed at the GitHub page linked below:  
- Github Page: [https://mas16.github.io/Tuition/TuitionSlides.html#1](https://mas16.github.io/Tuition/TuitionSlides.html#1)
